---
title: "Covid91 vaccine study Final2023"
author: "Tyler Frasier"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.




## Males
$H_$ The covid921 does not have an effect on the male population
$A_$ The covid921 does have an effect on the male population


```{r}
library(dplyr)
Males = filter(FinalData, FinalData$sex =='M')
```

##Methods 

Male is categorical, Covid921 vaccine Administration is explanatory variable. covid921 effectivness is a response variable, therefore we use cat~cat.



##Discriptive Results


##Graphical Results

```{r}
barchartGC(~infected+ treatment,data=Males)
barchartGC(~infected+ treatment,data=Males, type="percent")

```

This graphs shows that people who got the placebo had a higher percentage of getting covid 19 then oppose the people who didnt. The people whho received  the vaccine had a higher ok rate after the shot.


##Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```


##Inferential Results


## Females

$H_$ The covid921 does not have an effect on the Female population
$A_$ The covid921 does have an effect on the Female population


```{r}
library(dplyr)
Females = filter(FinalData, FinalData$sex =='F')
```

#Methods

Females is categorical, Covid921 vaccine Administration is explanatory variable. covid921 effectivness is a response variable, therefore we use cat~cat.


##Discriptive Results


##Graphical Results

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")

```

This graphs shows that people who got the placebo had a higher percentage of getting covid 19 then oppose the people who didnt. The people who recieved the vaccine had a balanced ok rate after the shot.


##Numerical Results 

```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```

##Inferential Results


## LGBTQ

$H_$ The covid921 does not have an effect on the LGBTQ population
$A_$ The covid921 does have an effect on the LGBTQ population

```{r}
library(dplyr)
LGBTQ = filter(FinalData, FinalData$LGBTQ =='gay')
```

#Methhods

LGBTQ is categorical, Covid921 vaccine Administration is explanatory variable. covid921 effectivness is a response variable, therefore we use cat~cat.

##Discriptive Results


##Graphical Results

```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")

```

This graphs shows that people who got the placebo had a lower percentage of getting covid 19 then oppose the people who didnt. The people whho recieved the vaccine had a balanced  ok rate after the shhot.


##Numerical Results 


```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```

##Inferential Results




## Druggies

$H_$ The covid921 does not have an effect on the Druggies population
$A_$ The covid921 does have an effect on the Druggies population


```{r}
library(dplyr)
Druggies = filter(FinalData, FinalData$DrugUser =='yes')
```

#Methods 

Druggies is categorical, Covid921 vaccine Administration is explanatory variable. Covid921 effectivness is a response variable, therefore we use cat~cat.

##Discriptive Results


##Graphical Results

```{r}
barchartGC(~infected + treatment,data=Druggies)
barchartGC(~infected + treatment,data=Druggies, type="percent")

```

This graphs shows that people who got the placebo had a lower percentage of getting covid 19 then oppose the people who didnt. The people whho recieved the vaccine had a lower ok rate after the shhot.



##Numerical Results 


```{r}
table1 <- xtabs(~infected + treatment, data=Druggies)
rowPerc(table1)
colPerc(table1)
```


##Inferential Results





# Overall Results and Conclusions


