---
title: "Covid91 vaccine study Final2023"
author: "Tyler Frasier"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.




## Males
$H_$ The covid921 does not have an effect on the male population
$A_$ The covid921 does have an effect on the male population


```{r}
library(dplyr)
Males = filter(FinalData, FinalData$sex =='M')
```

##Methods 

Male is categorical, Covid921 vaccine Administration is explanatory variable. covid921 effectivness is a response variable, therefore we use cat~cat.



##Discriptive Results


##Graphical Results

```{r}
barchartGC(~infected+ treatment,data=Males)
barchartGC(~infected+ treatment,data=Males, type="percent")
```

This graphs shows that people who got the placebo had a higher percentage of getting covid 19 then oppose the people who didnt. The people whho received  the vaccine had a higher ok rate after the shot.


##Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```
This graphh shows the people who had took the vaccine, 98.85% was ok, and people who took the placebo 97.45% was ok. The people without taking the placebo who took the vaccine 50% was ok and the people who took thhe vaccine, 49.52% was ok.


## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

the p value is 0.000000000001098. We reject thhe null hypothesis. The fisher exact is 0.00000000000004631.

##  Females

$H_$ The covid921 does not have an effect on the Female population
$A_$ The covid921 does have an effect on the Female population


```{r}
library(dplyr)
Females = filter(FinalData, FinalData$sex =='F')
```

# Methods

Females is categorical, Covid921 vaccine Administration is explanatory variable. covid921 effectivness is a response variable, therefore we use cat~cat.


## Discriptive Results


## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")

```

This graphs shows that people who got the placebo had a higher percentage of getting covid 19 then oppose the people who didnt. The people who recieved the vaccine had a balanced ok rate after the shot.


## Numerical Results 

```{r}
table2 <- xtabs(~infected + treatment, data=Females)
rowPerc(table2)
colPerc(table2)
```


This graph shows the people who had took the vaccine, 49.88% was ok, and people who took the placebo 50.12% was ok. The people without taking the placebo who took the vaccine 97.96% was ok and the people who took thhe vaccine, 98.63% was ok

## Inferential Results


```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

the p value is 0.00002732. we reject thhe null hypothesis. the fisher exact test is 0.0002162.

## LGBTQ

$H_$ The covid921 does not have an effect on the LGBTQ population
$A_$ The covid921 does have an effect on the LGBTQ population

```{r}
library(dplyr)
LGBTQ = filter(FinalData, FinalData$LGBTQ =='gay')
```

# Methhods

LGBTQ is categorical, Covid921 vaccine Administration is explanatory variable. covid921 effectivness is a response variable, therefore we use cat~cat.

## Discriptive Results


##Graphical Results

```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")

```

This graphs shows that people who got the placebo had a lower percentage of getting covid 19 then oppose the people who didnt. The people whho recieved the vaccine had a balanced  ok rate after the shhot.


##Numerical Results 


```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table3)
colPerc(table3)
```

This graph shows the people who had took the vaccine, 49.98% was ok, and people who took the placebo 50.02% was ok. The people without taking the placebo who took the vaccine 99.34% was ok and the people who took thhe vaccine, 96.46% was ok.



##Inferential Results


```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

the p value is 0.00006668. therefore we reject the null hypothesis. the fisher exact test p value is 0.00001029.

## Druggies

$H_$ The covid921 does not have an effect on the Druggies population
$A_$ The covid921 does have an effect on the Druggies population


```{r}
library(dplyr)
Druggies = filter(FinalData, FinalData$DrugUser =='yes')
```

#Methods 

Druggies is categorical, Covid921 vaccine Administration is explanatory variable. Covid921 effectivness is a response variable, therefore we use cat~cat.

##Discriptive Results


##Graphical Results

```{r}
barchartGC(~infected + treatment,data=Druggies)
barchartGC(~infected + treatment,data=Druggies, type="percent")

```

This graphs shows that people who got the placebo had a lower percentage of getting covid 19 then oppose the people who didnt. The people whho recieved the vaccine had a lower ok rate after the shhot.



##Numerical Results 


```{r}
table4 <- xtabs(~infected + treatment, data=Druggies)
rowPerc(table4)
colPerc(table4)
```

This graphh shows the people who had took the vaccine,48.81 % was ok, and people who took the placebo 51.19% was ok. The people without taking the placebo who took the vaccine 99.6% was ok and the people who took thhe vaccine, 96.94% was ok.



##Inferential Results


```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```


the p value is 0.00002842. therefore we reject the null hypothhesis. the fisher exact test p value is 0.0000232.

# Overall Results and Conclusions
in conclusions the original is, the vaccine is suspected that it may exhibit different effectiveness across the variety of subpopulations tracked? thhere is a difference in the effectiveness across the variety of subpopulations. the males and females who took the vaccine did better than the others who took the placebo. the lgbtq and druggies who took the vaccine didnt do better than the others whho took the placebo. the male p value 0.0001098. thhe female p value was 0.00002732. the lgbtq p value 0.00006668. the druggies p value is 0.0002842. sice all the p value is smaller than 0.05 we reject the null hypothesis.

